---
title: "Lesson - 5. 16.05.2023"
author: "Alessandro Bramucci"
output: pdf_document
date: "16.05.2023"
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Frequency distribution^[Exercise taken from Spiegel, M.R. Stephens, L.J. Statistics (Schaum's Outline Series).]

The final grade in Mathematics of 80 students are recorded as follows:

```{r}
x <- c(68, 84, 75, 82, 68, 90, 62, 88, 76, 93,
	73, 79, 88, 73, 60, 93, 71, 59, 85, 75,
	61, 65, 75, 87, 74, 62, 95, 78, 63, 72,
	66, 78, 82, 75, 94, 77, 69, 74, 68, 60,
	96, 78, 89, 61, 75, 95, 60, 79, 83, 71,
	79, 62, 67, 97, 78, 85, 76, 65, 71, 75,
	65, 80, 73, 57, 88, 78, 62, 76, 53, 74,
	86, 67, 73, 81, 72, 63, 76, 75, 85, 77)
```

\begin{gather}
	68, 84, 75, 82, 68, 90, 62, 88, 76, 93, \nonumber \\
	73, 79, 88, 73, 60, 93, 71, 59, 85, 75, \nonumber \\
	61, 65, 75, 87, 74, 62, 95, 78, 63, 72, \nonumber \\
	66, 78, 82, 75, 94, 77, 69, 74, 68, 60, \nonumber \\
	96, 78, 89, 61, 75, 95, 60, 79, 83, 71, \nonumber \\
	79, 62, 67, 97, 78, 85, 76, 65, 71, 75, \nonumber \\
	65, 80, 73, 57, 88, 78, 62, 76, 53, 74, \nonumber \\
	86, 67, 73, 81, 72, 63, 76, 75, 85, 77 \nonumber
\end{gather}

1) The highest grade.

```{r}
max(x)
```

2) The lowest grade.

```{r}
min(x)
```

3) The range.

```{r}
max(x) - min(x)
```

4) The grades of the five highest ranking students.

```{r}
x_sorted <- sort(x)
x_sorted[(length(x_sorted)-4):length(x_sorted)]
```

5) The grades of the five lowest ranking students.

```{r}
x_sorted[1:5]
```

6) The grade of the tenth high ranking student.

```{r}
x_sorted[length(x_sorted)-9]
```

7) The number of students who received grades of 75 or higher.

```{r}
sum(x >= 75)
```

8) The number of students who received grades below 85.

```{r}
sum(x < 85)
```

9) The percentage of students who received grades higher than 65 but not higher than 85.

```{r}
sum(x > 65 & x <= 85) / length(x) * 100
```

10) Produce an histogram of the data.

```{r}
hist(x,
     ylim = c(0, 20),
     main = "Final grades in Mathematics (N = 80)",
     xlab = "Grades")
```
