---
title: 'Lesson 14'
subtitle: 'Data visualisation with R'
author: "Alessandro Bramucci"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
runtime: shiny
output: 
  bookdown::html_document2:
---

```{r setup, include=FALSE}
library(shiny)
library(dplyr)
library(plotly)

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Embedded Shiny apps

## Inline application

Modified version of the k-means clustering Shiny app using the *iris* dataset.

```{r, echo=FALSE}

vars <- setdiff(names(iris), "Species")

shinyApp(

pageWithSidebar(
  
  headerPanel('Iris k-means clustering'),
  
  sidebarPanel(
    selectInput('xcol', 'X Variable', choices = vars, selected = "Sepal.Width"), 
    selectInput('ycol', 'Y Variable', choices = vars, selected = "Sepal.Length"),
      checkboxInput("centers", "Select number of clusters", FALSE),
    conditionalPanel(condition = "input.centers", 
      sliderInput('clusters', 'Cluster count', 1, min = 1, max = 9)
    )
  ),
  
  mainPanel(
    plotlyOutput('plot1')
  )
  
),

server = function(input, output, session) {
  
  observe({
    vars_x <- setdiff(names(iris), c("Species", input$xcol))
    updateSelectInput(session, 'ycol', choices = vars_x, selected = input$ycol)
  })
  
  observe({
    vars_y <- setdiff(names(iris), c("Species", input$ycol))
    updateSelectInput(session, 'xcol', choices = vars_y, selected = input$xcol)
  })

  selectedData <- reactive({
    iris[, c(input$xcol, input$ycol)]
  })

  clusters <- reactive({
    kmeans(selectedData(), input$clusters)
  })
  
  my_colors <- reactive({
    c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
      "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999")[1:input$clusters]
  })

  output$plot1 <- renderPlotly({
    
    p <- plot_ly() %>% 
      add_trace(x = selectedData()[[1]],
                y = selectedData()[[2]],
                type = "scatter",
                mode = "markers",
                colors = my_colors(),
                color = as.factor(clusters()$cluster),
                marker = list(size = 10)) %>% 
  layout(xaxis = list(title = input$xcol,
                      dtick = 0.5),
         yaxis = list(title = input$ycol,
                      dtick = 0.5),
         showlegend = FALSE) %>%
  config(displayModeBar = F)

    if (input$centers == TRUE) {
      
      p <- p %>%
        add_trace(x = clusters()$centers[,1],
                  y = clusters()$centers[,2],
                  type = "scatter",
                  mode = "markers",
                  name = "Center",
                  marker = list(symbol = "x",
                                size = 20,
                                color = "black"))
      
    }
    
    p
    
  })

},

  options = list(height = 500)

)

```

## External application

Original version of the k-means clustering Shiny app developed by [Carson Sievert](https://github.com/cpsievert).

```{r, out.width = 900}

knitr::include_app("https://gallery.shinyapps.io/050-kmeans-example/", height = 500)

```

# Useful links and resources

The basics of the k-means clustering:

- https://uc-r.github.io/kmeans_clustering

K-means example in Shiny:

- https://shiny.posit.co/r/gallery/start-simple/kmeans-example/

How to embedd a Shiny app in an rmarkdown / bookdown document:

- https://bookdown.org/yihui/rmarkdown/shiny-embedded.html

- https://bookdown.org/yihui/bookdown/web-pages-and-shiny-apps.html
