---
title: 'Lesson 10'
subtitle: 'Data visualisation with R'
author: "Alessandro Bramucci"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 3
bibliography: references.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(gapminder)
library(stargazer)
library(kableExtra)

```

# The most populous countries in 2007

Table \@ref(tab:table1) reports the three most populous countries in 2007 broken down by continent.

```{r table1}

pop_countries <- gapminder %>% 
  filter(year == 2007) %>% 
  group_by(continent) %>%
  select(country, pop) %>% 
  arrange(desc(pop)) %>% 
  slice(1:3) %>% 
  ungroup()

pop_table <- pop_countries %>%
  select(country, pop) %>% 
  kbl(booktabs = TRUE,
      format.args = list(decimal.mark = ',',
                         big.mark = "\'"),
      caption = "The three most populous countries by continent in 2007. Source: Gapminder.",
      col.names = c("Country",
                    "Population")) %>% 
  pack_rows("Africa", 1, 3) %>%
  pack_rows("Americas", 4, 6) %>%
  pack_rows("Asia", 7, 9) %>%
  pack_rows("Europe", 10, 12) %>%
  pack_rows("Oceania", 13, 14) %>%
  kable_styling(latex_options = c("hold_position"))

pop_table

```

# Life expectancy and per capita GDP

From Figure \@ref(fig:figure) we can observe the existence of a positive relationship between per capita GDP growth and life expectancy.

```{r figure, fig.align = "center", out.width = '80%', out.height = '30%', fig.cap = "Relationship between per capita GDP growth and change in life expectancy, 1952-2007."}

df_graph <- gapminder %>% 
  group_by(country) %>% 
  mutate(pcGDPgr = c(NA, diff(gdpPercap))/lag(gdpPercap) * 100,
         dlifeExp = c(NA, diff(lifeExp))) %>% 
  drop_na()

p <- ggplot(df_graph, aes(y = dlifeExp,
                          x = pcGDPgr)) +
  geom_point(aes(color = continent)) + 
  geom_smooth(method = lm, se = FALSE) + 
  scale_color_manual(values = continent_colors) + 
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(color = NULL,
       y = "Change in life expectancy",
       x = "Per capita GDP growth",
       caption = "Source: Gapminder.")

p

```

## Regression analysis

This document is written using the package *bookdown* [@bookdown]. Data are obtained from the *Gapminder* dataset [@gapminder].

### Per capita GDP growth and life expectancy

```{r, results = 'asis'}

reg <- lm(dlifeExp ~ pcGDPgr, data = df_graph)

# Regression table
stargazer(reg,
          header = FALSE,
          type = "latex",
          label = "tab:regressions",
          title = "Relationship between per capita GDP growth and life expectancy, 1952-2007.",
          dep.var.labels = c("dlifeExp"),
          covariate.labels = c("pcGDPgr"),
          keep.stat = c("n", "rsq"))

```

### Inclusion of dummy variables

In econometrics, a dummy variable is a binary variable used to represent categorical or qualitative variables in regression analysis and to capture the effects of different categories on the dependent variable. Dummy variables are created by assigning a value of 1 to one category or group, while all other categories or groups are assigned a value of 0. Since we have a variable called "continent" with five categories, i.e. the five continents, we can create a dummy variable called "Europe" and assign a value of 1 to countries in Europe and 0 to countries in other continents. 

By including dummy variables in regression models, we can estimate the effect of different categories on the outcome variable.^[There are several ways in R to create dummy variables.] When we want to include dummy variables in a regression, we must avoid the so-called dummy variables trap. What is the dummy variables trap? The dummy variable trap arises when we include dummy variables for all categories of a categorical variable in a regression model, for example all continents in our case. Including dummy variables for all categories creates a linear dependency among the variables, causing perfect multicollinearity.

To understand why this happens, consider a categorical variable with *k* categories like our five continents, for example. If we create *k-1* dummy variables, where one category is excluded as the reference, the information about the excluded category is captured in the constant term of the regression equation. We must then interpret the coefficient of the dummy variable with respect to the constant term, i.e. the benchmark category. In Table \@ref(tab:regressions) the benchmark category is set to "Europe".^[Here we have used the function *relevel(continent, ref = "Europe")*.] If different dummy variable sets are included, such as for example for continents but also whether the country is located north or south of the hemisphere, the interpretation of the constant as a reference category will no longer be possible.

```{r, include = FALSE}

library(modeldb)

df_graph_dummy <- gapminder %>% 
  add_dummy_variables(continent, auto_values = TRUE)

```

```{r, results = 'asis'}

reg3 <- lm(dlifeExp ~ pcGDPgr + relevel(continent, ref = "Europe"), data = df_graph)

reg4 <- lm(lifeExp ~ pcGDPgr + relevel(continent, ref = "Europe"), data = df_graph)

stargazer(reg3,
          reg4,
          header = FALSE,
          type = "latex",
          label = "tab:regressions",
          title = "Relationship between per capita GDP growth and life expectancy including continent dummy variables, 1952-2007.",
          dep.var.labels = c("dlifeExp", "lifeExp"),
          covariate.labels = c("pcGDPgr", "Americas",
                               "Asia", "Africa", "Oceania"),
          keep.stat = c("n", "rsq"))

```

\newpage

# References
