---
title: 'Lesson 8'
subtitle: 'Data visualisation with R'
author: "Alessandro Bramucci"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(dplyr)
library(ggplot2)
library(gapminder)
library(kableExtra)

```

# Exploring the data

```{r, include=FALSE}

length(unique(gapminder$country))

n_countries <- gapminder %>%
  filter(year == 2007) %>%
  summarise(n_country = n())

```

How many countries in the Gapminder dataset? In the Gapminder dataset there are `r n_countries` countries. How many countries per continent?

```{r, echo = FALSE}
gapminder %>% 
  filter(year == 2007) %>% 
  group_by(continent) %>% 
  summarise(`Nr. country` = n()) %>% 
  kbl(booktabs = T) %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```

To customise your table using *KableExtra* take a look [here](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) and [here](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html).

# Visualise the data

Is there a positive relationship between GDP per capita and life expectancy?

```{r figure1, echo = FALSE, fig.align = "center", 	out.width = '80%',out.height = '30%', fig.cap = "Relationship between income per person and life expectancy, year 2007."}

gapminder %>% 
  filter(year == 2007) %>%
  mutate(ln_gdpPercap = log(gdpPercap)) %>% 
  ggplot(aes(x = ln_gdpPercap, 
             y = lifeExp)) +
  geom_point(aes(color = continent)) + 
  geom_smooth(method = lm, se = FALSE) + 
  theme_bw() + 
    labs(color = NULL,
         y = "Life expectancy at birth",
         x = "Log GDP per capita (in $)",
         caption = "Source: Gapminder.")

```

Figure 1 shows the existence of a positive relationship between log per capita GDP and life expectancy.

<!-- To cross-reference within the text use the bookdown output format in the YAML output field -->
<!-- bookdown::pdf_document2: default.  -->
<!-- Take a look here https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html)  -->
