---
title: 'Lesson 9 - Homework'
subtitle: 'Data visualisation with R'
author: "Alessandro Bramucci"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(gapminder)
library(kableExtra)

```

# World population by continent in 2007

## Table

Table \@ref(tab:table) reports population data by continent in 2007 as a percentage of the world's population.

```{r, table}

world_pop <- gapminder %>%
  filter(year == 2007) %>%
  summarise(w_pop = sum(pop))

df_2007_pc <- gapminder %>% 
  filter(year == 2007) %>%
  group_by(continent) %>% 
  summarise(`2007` = round(sum(pop)/as.numeric(world_pop) * 100, digits = 2))

# For later use we extract the value for Asia
df_2007_pc_Asia <- df_2007_pc %>% 
  filter(continent == "Asia")
  
table_2007 <- df_2007_pc %>% 
  kbl(booktabs = TRUE,
      align = "lr",
      caption = "World population by continent, year 2007. Source: Gapminder.",
      col.names = c("Continent",
                    "Population (in %)")) %>% 
  kable_styling(latex_options = c("hold_position"))

table_2007
  
```

\newpage

## Pie chart

```{r piechart, fig.cap = "World population by continent, year 2007. Source: Gapminder.", fig.align = "center", out.width = '80%', out.height = '30%'}

# No need to calculate percentage values
df_2007 <- gapminder %>% 
  filter(year == 2007) %>% 
  group_by(continent) %>% 
  summarise(`2007` = sum(pop))

ggplot(df_2007, aes(x = "x", 
                    y = `2007`,
                    fill = continent)) +
        geom_bar(stat = "identity", 
                 width = 1, 
                 color = "white") +
        coord_polar("y", 
                    start = 0) + 
        theme_void() +
  scale_fill_manual(values = continent_colors) + 
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(title = NULL))

```

From Figure \@ref(fig:piechart) we can see that Asia is the most populous continent. In 2007, the percentage of the population of the Asian continent represents `r paste0(as.numeric(df_2007_pc_Asia[2]),"%")` of the world's population.
